---
title: "prod2"
author: "Ido Klein"
date: "3/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## packages and definations

piping and data manipulation - {tidyverse}\
geometry manipulation - {sf}, {lwgeom}, {nngeo}\
network analysis - {sfnetworks}, {tidygraph}\
gtfs loading - {gtfstools}\
visualization - {mapview}\
hebrew encoding
```{r load libraries and setup workspace}
library(tidyverse)
library(sf)
library(lwgeom)
library(nngeo)
library(sfnetworks)
library(tidygraph)
library(gtfstools)
library(mapview)
Sys.setlocale(locale = "hebrew")
```

## loading GTFS, filtering Beer Sheva

```{r load gtfs}
spo_gtfs <- read_gtfs("C:/Users/yehuda/Downloads/gtfs.zip",encoding = "UTF-8")
bs_pol <- spo_gtfs$routes %>% 
  # at first, only for Dan Beer sheva
  filter(agency_id == 32) %>% 
  select(route_id,route_desc) %>% 
  # join with trips
  left_join(spo_gtfs$trips, by = "route_id") %>% 
  left_join(spo_gtfs$stop_times, by = "trip_id") %>% 
  distinct(stop_id) %>% 
  left_join(spo_gtfs$stops, by = "stop_id") %>% 
  st_as_sf(coords = c("stop_lon","stop_lat"),crs = 4326) %>% 
  summarise() %>% 
  st_convex_hull() 

bs_gtfs <- filter_by_sf(spo_gtfs,bs_pol)
```

## create building blocks

```{r building clocks}
bs_lines <- get_trip_geometry(bs_gtfs) %>% 
  # join trips to their geometry
  left_join(bs_gtfs$trips, by = "trip_id") %>% 
  # get only one geometry per shape_id
  select(shape_id) %>%
  group_by(shape_id) %>% 
  mutate(rn = row_number()) %>% 
  filter(rn == 1) %>%
  ungroup() %>% 
  select(-rn) %>% 
  st_sf() %>% 
  st_transform(2039) 
# create layer for each stop
bs_stops <- bs_gtfs$stops %>% 
  st_as_sf(coords = c("stop_lon","stop_lat"),crs = 4326) %>% 
  st_transform(2039) %>% 
  select(stop_id)
# create layer for each intersection
intersections <- 
  # explode all lines
  st_segments(bs_lines$geometry) %>% 
  st_as_sf() %>% 
  # get only distinct segments
  distinct()  %>% 
  as_sfnetwork(directed = F) %>% 
  # subdivide segments that intersect (assuming planar graph)
  convert(to_spatial_subdivision) %>% 
  # smooth useless nodes
  convert(to_spatial_smooth) %>%
  mutate(inter_id = paste0("inter_",row_number()),
         degree = centrality_degree(mode = "all")) %>% 
  st_as_sf() 

# for each stop, join a relevant line
joined <- bs_stops %>% 
  st_join(bs_lines %>% 
            st_buffer(-10,singleSide = T) ,left = FALSE) %>% 
  arrange(stop_id)
# for each intesection, join a relevant line
joined_inter <- intersections %>% 
  st_join(bs_lines %>% 
            st_buffer(-10,singleSide = T) ,left = FALSE)
```

## processing each shape_id

```{r processing shape_ids}
res <- map(1:nrow(bs_lines),function(.x){
  print(bs_lines[.x,]$shape_id)
  # filter relevant stops and intersections
  one <- joined %>% filter(shape_id == bs_lines[.x,]$shape_id)
  two <- joined_inter%>% filter(shape_id == bs_lines[.x,]$shape_id)
  # blend intersections to sfnetwork
  net <- bs_lines[.x,] %>% 
    as_sfnetwork(directed = T) %>% 
    convert(to_spatial_subdivision) %>%
    st_network_blend(two)
  # filter segments where there are stops
  rel_rows <- net %E>%
    st_as_sf() %>% 
    mutate(rn = row_number()) %>% 
    st_buffer(-10,singleSide = T) %>% 
    filter(apply(st_intersects(geometry,one,sparse = F),1,any)) 
  # get stops ids in order to blenf into network
  rel_idx <- net %E>%
    st_as_sf() %>% 
    mutate(rn = row_number()) %>% 
    st_buffer(-10,singleSide = T) %>% 
    st_intersects(one,sparse = F) %>% 
    apply( 1,which)
  # create sfnetwork for each segemnt
  nets <- net %E>%
    st_as_sf() %>% 
    mutate(rn = row_number(),
           idx = rel_idx) %>% 
    filter(rn %in% rel_rows$rn) %>% 
    mutate(
      # create an sfnetwork for each segment according to row number
      net1 = map2(geometry,rn, ~.x %>% 
                         st_sfc(crs = 2039) %>% 
                         st_sf() %>% 
                         as_sfnetwork(directed = T) %E>% 
                         mutate(rn = .y)),
           # blend stops into each sfnetwork
           net2 = map2(net1,idx, function(x,y){
             st_network_blend(x,one[y,]) %>% 
               mutate(rn2 = row_number())
           })) %>% 
    pull(net2)
  # get all joined net stops
  stops <- bind_rows(map(nets,~.x %N>% st_as_sf())) %>% 
    filter(!is.na(stop_id)) %>% 
    distinct()
  # bind all graphs together, including those segments without stops
  final_net <- net %E>%
    mutate(rn = row_number()) %>%
    filter(!rn %in% rel_rows$rn) %E>%
    bind_graphs(bind_graphs(nets)) %>%
    # convert to sfnetwork
    as_sfnetwork(edges_as_lines = T,directed = T) %E>% 
    st_as_sf() %>%
    # arrange by order
    arrange(rn,rn2) %>% 
    as_sfnetwork(directed = T)
  route <- final_net %E>% 
    st_as_sf() %>% 
    mutate(cs = cumsum(st_length(geometry))) 
  # create cuts
  cuts <- final_net %>% 
    st_join(stops) %>%
    st_join(two) %>% 
    mutate(rn = row_number()) %>% 
    as_tibble() %>% 
    as_tibble() %>%
    select(geometry,stop_id,inter_id,rn)
  # process route and cuts
  phase_1 <- route %>%
    # join all cuts
    left_join(cuts,by = c("from" = "rn")) %>%
    left_join(cuts,by = c("to" = "rn")) %>%
    # fill all shape_ids to prevent errors
    fill(shape_id,.direction = "downup") %>%
    # give each origin and destination an id: stop id if stop, intersection id if intersection, start\end if edges
    mutate(origin = ifelse(!is.na(stop_id.x),stop_id.x,ifelse(!is.na(inter_id.x),inter_id.x,NA)),
           destination = ifelse(!is.na(stop_id.y),stop_id.y,ifelse(!is.na(inter_id.y),inter_id.y,NA)),
           origin = ifelse(row_number() == 1,paste0("start_",shape_id),origin),
           destination = ifelse(row_number() == nrow(.),paste0("end_",shape_id),destination)) %>%
    select(origin,destination,from,to) %>%
    fill(origin) %>%
    fill(destination,.direction = "up") %>%
    # preserve order
    mutate(lag1 = lag(origin),
           lag2 = lag(destination),
           counter =coalesce(ifelse(origin == lag1 & destination == lag2,0,1),1) %>% cumsum()) %>% 
    select(-lag1,-lag2)
  # join created lines, some have no named edge
  phase_2 <- phase_1 %>%
    group_by(origin,destination,counter) %>%
    summarise(do_union=T)
  # prevent bugs if no multilinestring
  phase_21 <- phase_2 %>%
    filter(st_geometry_type(geometry.x) == "MULTILINESTRING")
  if(nrow(phase_21) == 0){
    phase_3 <- phase_2
  }else{
    phase_3 <- bind_rows(phase_2 %>%
                           filter(!st_geometry_type(geometry.x) == "MULTILINESTRING"),
                         phase_2 %>%
                           filter(st_geometry_type(geometry.x) == "MULTILINESTRING") %>%
                           st_line_merge())
  }
  phase_3 %>%
    ungroup() %>%
    arrange(counter) %>%
    # st_buffer(-10,singleSide = T) %>%
    mutate(rn3 = row_number())
})
```

## map of products
```{r map of lines and cuts}
names(res) <- bs_lines$shape_id
with_shapes <- map(names(res),~res[[.x]] %>% mutate(shape_id = .x)) %>% bind_rows()
starts <- with_shapes %>% st_startpoint() %>% st_sf()
m1 <- mapview(with_shapes) + mapview(starts)
m1@map
```

## stops in routes - segmenting 1,2,3 order cuts
```{r order cuts}
stops_in_routes <- bs_gtfs$routes %>% 
  left_join(bs_gtfs$trips,by = "route_id") %>% 
  left_join(bs_gtfs$stop_times,by = "trip_id") %>% 
  select(route_id,route_desc,shape_id,stop_id,agency_id) %>% 
  distinct()  
# remove train - useless
to_remove <- stops_in_routes %>% filter(agency_id== 2) %>% distinct(shape_id) %>% pull(shape_id)
with_shapes_2 <- with_shapes %>% filter(!shape_id %in% to_remove)

with_shapes_3 <- with_shapes_2 %>% 
  # for each shape_id, add cumulative sum of length to know distance in line
  mutate(length = st_length(geometry.x)) %>% 
  as_tibble() %>% 
  group_by(shape_id) %>% 
  mutate(length= as.numeric(length),
         cs = replace_na(lag(cumsum(length)),0)) %>% 
  left_join(stops_in_routes,by = c("origin" = "stop_id","shape_id")) %>% 
  left_join(stops_in_routes,by = c("destination" = "stop_id","shape_id"))
# getting relevent makatim
makats <- bs_gtfs$routes %>% 
  filter(!agency_id== 2) %>% 
  left_join(bs_gtfs$trips,by = "route_id") %>% 
  select(route_id,route_desc,shape_id) %>% 
  distinct()

prod1 <- makats %>% 
  left_join(with_shapes_3,by = "shape_id") %>%
  select(-route_id,-route_id.x,-route_id.y,-agency_id.x,-agency_id.y,-rn3) %>% 
  mutate(start = str_detect(origin, "start"),
         end = str_detect(destination, "end"),
         is_stop_in_orig = coalesce(route_desc == route_desc.x,FALSE),
         is_stop_in_dest = coalesce(route_desc == route_desc.y,FALSE),
         is_orig_intersection = str_detect(origin,"inter"),
         is_dest_intersection = str_detect(destination,"inter"),
         orig_cut = ifelse(start | is_stop_in_orig, "first",
                           ifelse(is_orig_intersection,"third","second")),
         dest_cut = ifelse(end | is_stop_in_dest, "first",
                           ifelse(is_dest_intersection,"third","second")))
prod2 <- prod1 %>% 
  select(-c(route_desc.x:is_dest_intersection)) %>% 
  mutate(segment_type = ifelse(orig_cut == "third" | dest_cut == "third", "third",
                                   ifelse(orig_cut =="second"|dest_cut =="second","second","first")),
         counter_first = orig_cut == "first",
         counter_second = orig_cut == "second",
         counter_third = orig_cut == "third"
         ) %>% 
  group_by(shape_id) %>% 
  mutate(counter_seg_first = cumsum(counter_first)) %>% 
  ungroup() %>% 
  group_by(shape_id,counter_seg_first) %>% 
  mutate(counter_seg_second = cumsum(counter_second),
         has_second = sum(counter_second) > 0) %>% 
  ungroup() %>% 
  group_by(shape_id,counter_seg_first,counter_seg_second) %>% 
  mutate(counter_seg_third = cumsum(counter_third),
         has_third = sum(counter_third)>0) %>% 
  mutate(geometry.x = st_transform(geometry.x, 4326))
prod2
```
 
## creating tables

```{r to tables}
final <- list()
final$routes <- makats %>% 
  mutate(route_id=as.integer(route_id),
         shape_id = shape_id) %>% 
  select(route_id,shape_id,route_desc)
st_x = function(x) round(st_coordinates(x)[,1],5)
st_y = function(x) round(st_coordinates(x)[,2],5)
final$cuts <- bind_rows(
  prod2 %>% ungroup() %>% 
    select(origin,geometry.x) %>% 
    mutate(geom = st_startpoint(geometry.x)) %>%
    select(-geometry.x) %>% 
    st_sf() %>% 
    rename(cut_id = origin),
  prod2 %>% 
    ungroup() %>% 
    select(destination,geometry.x) %>% 
    mutate(geom = st_endpoint(geometry.x)) %>%
    select(-geometry.x) %>% 
    st_sf() %>% 
    rename(cut_id = destination)
) %>% distinct() %>% 
  mutate(intersection = str_detect(cut_id,"inter"),
         start = str_detect(cut_id,"start"),
         end = str_detect(cut_id,"end"),
         type = ifelse(intersection,"inter",ifelse(start,"start",ifelse(end,"end","stop"))),
         lon = st_x(geom),
         lat = st_y(geom)) %>% 
  st_drop_geometry() %>% 
  select(cut_id,type,lon,lat)

final$segments <- prod2 %>% 
  ungroup() %>% 
  select(geometry = geometry.x,length) %>% 
  distinct() %>% 
  mutate(seg_UID = row_number(),
         geometry = map_chr(geometry,st_as_text)) %>% 
  select(seg_UID,length,geometry)


tmp <- prod2 %>%
  left_join(final$segments %>% st_as_sf(wkt="geometry",crs = 4326), by = c("geometry.x"="geometry"))
final$first_order_cuts <- tmp %>% 
  ungroup() %>% 
  select(shape_id,
         first_order_seg_id = counter_seg_first,
         from_cut = origin,
         to_cut = destination) %>% 
         # seg_UID,
         # distance_from_origin = cs,
         # counter,
         # has_second,
         # has_third) %>% 
  # mutate(serial_cut_number = ifelse(has_second|has_third, NA_integer_,counter),
         # seg_UID = ifelse(has_second|has_third, NA_integer_,seg_UID)) %>% 
  # select(-counter,-has_second,-has_third) %>% 
  group_by(shape_id,first_order_seg_id) %>% 
  mutate(from_cut = first(from_cut),
            to_cut = last(to_cut),
            # distance_from_origin = first(distance_from_origin)
         ) %>% 
  distinct()
final$second_order_cuts <- tmp %>% 
  ungroup() %>% 
  filter(has_second) %>% 
  select(shape_id,
         first_order_seg_id = counter_seg_first,
         second_order_seg_id = counter_seg_second,
         from_cut = origin,
         to_cut = destination) %>% 
         # seg_UID,
         # distance_from_origin = cs,
         # counter,
         # has_third) %>% 
  # mutate(serial_cut_number = ifelse(has_third, NA_integer_,counter),
         # seg_UID = ifelse(has_third, NA_integer_,seg_UID)) %>% 
  # select(-counter,-has_third) %>% 
  distinct() %>% 
  group_by(shape_id,first_order_seg_id,second_order_seg_id) %>% 
  mutate(from_cut = first(from_cut),
            to_cut = last(to_cut),
            # distance_from_origin = first(distance_from_origin)
         ) %>% 
  distinct()  
final$third_order_cuts <- tmp %>% 
  ungroup() %>% 
  filter(has_third) %>% 
  select(shape_id,
         first_order_seg_id = counter_seg_first,
         second_order_seg_id = counter_seg_second,
         third_order_seg_id = counter_seg_third,
         from_cut = origin,
         to_cut = destination,

         # seg_UID,
         # distance_from_origin = cs,
         # serial_cut_number =counter,
         has_second) %>% 
  mutate(second_order_seg_id = ifelse(has_second,second_order_seg_id,NA_integer_)) %>% 
  select(-has_second) 
final$route_order <- tmp %>% 
  ungroup() %>% 
  select(shape_id, 
         serial_cut_number = counter, 
         first_order_seg_id= counter_seg_first,
         second_order_seg_id= counter_seg_second,
         third_order_seg_id= counter_seg_third,
         seg_UID,
         distance_from_origin = cs,
         serial_cut_number =counter,
         has_second,
         has_third) %>% 
  mutate(second_order_seg_id = ifelse(has_second,second_order_seg_id,NA_integer_),
         third_order_seg_id = ifelse(has_third,third_order_seg_id,NA_integer_)) %>% 
  select(-has_second,-has_third)
map(names(final),~write_csv(final[[.x]],paste0(.x,".csv")))
```

```{r}
 new_model <- df_to_model(
     list=names(final), envir=as.environment(final)
)
new_model %>% 
  add_foreign_key("route_order","seg_UID","segments","seg_UID") %>%
  add_foreign_key("first_order_cuts","from_cut","cuts","cut_id") %>% 
  add_foreign_key("first_order_cuts","to_cut","cuts","cut_id") %>% 
  add_foreign_key("second_order_cuts","from_cut","cuts","cut_id") %>% 
  add_foreign_key("second_order_cuts","to_cut","cuts","cut_id") %>%
  add_foreign_key("third_order_cuts","from_cut","cuts","cut_id") %>% 
  add_foreign_key("third_order_cuts","to_cut","cuts","cut_id") %>% 
  add_foreign_key("first_order_cuts","first_order_seg_id","route_order","first_order_seg_id") %>% 
  add_foreign_key("second_order_cuts",c("first_order_seg_id","second_order_seg_id"),"route_order",c("first_order_seg_id","second_order_seg_id")) %>% 
  add_foreign_key("third_order_cuts",c("first_order_seg_id","second_order_seg_id","third_order_seg_id"),"route_order",c("first_order_seg_id","second_order_seg_id","third_order_seg_id")) %>% 
  add_foreign_key("route_order","shape_id","routes","shape_id") %>%
  add_foreign_key("first_order_cuts","shape_id","routes","shape_id") %>%
  add_foreign_key("second_order_cuts","shape_id","routes","shape_id") %>%
  add_foreign_key("third_order_cuts","shape_id","routes","shape_id") %>%
  model_relational_data()
```

