---
title: "prod2"
author: "Ido Klein"
date: "3/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## packages and definations

piping and data manipulation - {tidyverse}\
geometry manipulation - {sf}, {lwgeom}, {nngeo}\
network analysis - {sfnetworks}, {tidygraph}\
gtfs loading - {gtfstools}\
visualization - {mapview}\
hebrew encoding
```{r load libraries and setup workspace, message=FALSE, warning=FALSE,results = FALSE}
library(tidyverse)
library(sf)
library(lwgeom)
library(nngeo)
library(sfnetworks)
library(tidygraph)
library(gtfstools)
library(mapview)
Sys.setlocale(locale = "hebrew")
```

## loading GTFS, filtering Beer Sheva
first, we filter the Israeli gtfs file from 7.7.2021 to only include the lines passing through beersheva
```{r load gtfs, message=FALSE, warning=FALSE,results = FALSE}
spo_gtfs <- read_gtfs("C:/Users/yehuda/Downloads/gtfs.zip",encoding = "UTF-8")
# bs_pol <- spo_gtfs$routes %>% 
#   # at first, only for Dan Beer sheva
#   filter(agency_id == 32) %>% 
#   select(route_id,route_desc) %>% 
#   # join with trips
#   left_join(spo_gtfs$trips, by = "route_id") %>% 
#   left_join(spo_gtfs$stop_times, by = "trip_id") %>% 
#   distinct(stop_id) %>% 
#   left_join(spo_gtfs$stops, by = "stop_id") %>% 
#   st_as_sf(coords = c("stop_lon","stop_lat"),crs = 4326) %>% 
#   summarise() %>% 
#   st_convex_hull() 

# bs_gtfs <- filter_by_sf(spo_gtfs,bs_pol)
bs_gtfs <- spo_gtfs
```

## create building blocks
using the gtfs elements, we create three building block for the analysis: 
1. bs_lines - line geometries of the shapes of each shape_id  
2. bs_stops - point geometries of all stops
3. intersections - point geometries of all the times bs_lines intesect with each other
the creation of the first two is quite trivial. the intersections are created the following way:
- exploding all lines  
- finding only distinct segments  
- turning them into a spatial graph
- subdividing the network by finding shared implicit nodes inside edges
- un-exploding the network by smoothing the edges (removing 2-degree nodes)
we have 2 helper tables:
1. joined - joining all shape_ids for each stop, by finding which of them are 10 meters to the right of the line
1. joined_inter - joining all shape_ids for each intersection, by finding which of them are 10 meters to the right of the line
```{r building clocks, message=FALSE, warning=FALSE,results = FALSE}
bs_lines <- get_trip_geometry(bs_gtfs) %>% 
  # join trips to their geometry
  left_join(bs_gtfs$trips, by = "trip_id") %>% 
  # get only one geometry per shape_id
  select(shape_id) %>%
  group_by(shape_id) %>% 
  mutate(rn = row_number()) %>% 
  filter(rn == 1) %>%
  ungroup() %>% 
  select(-rn) %>% 
  st_sf() %>% 
  st_transform(2039) 
# create layer for each stop
bs_stops <- bs_gtfs$stops %>% 
  st_as_sf(coords = c("stop_lon","stop_lat"),crs = 4326) %>% 
  st_transform(2039) %>% 
  select(stop_id)
# create layer for each intersection
intersections <- 
  # explode all lines
  st_segments(bs_lines$geometry) %>% 
  st_as_sf() %>% 
  # get only distinct segments
  distinct()  %>% 
  as_sfnetwork(directed = F) %>% 
  # subdivide segments that intersect (assuming planar graph)
  convert(to_spatial_subdivision) %>% 
  # smooth useless nodes
  convert(to_spatial_smooth) %>%
  mutate(inter_id = paste0("inter_",row_number()),
         degree = centrality_degree(mode = "all")) %>% 
  st_as_sf() 

# for each stop, join a relevant line
joined <- bs_stops %>% 
  st_join(bs_lines %>% 
            st_buffer(-10,singleSide = T) ,left = FALSE) %>% 
  arrange(stop_id)
# for each intesection, join a relevant line
joined_inter <- intersections %>% 
  st_join(bs_lines %>% 
            st_buffer(-10,singleSide = T) ,left = FALSE)
```

## processing each shape_id
this is the main workhorse.  
in this process, 
we take each shape_id  
turn it into a spatial network  
subdividing the network to include self intersections  
blending the intersections into the network  
inserting stops into the relevant edges and creating multiple sfnetworks  
binding all graphs together in a way that keep the order of the nodes according to their location in the line
adding a cumulative sum of the length of the segments to get the distance from origin  
adding stop and intersection ids  
adding start point and end point ids  
removing those nodes that are internal and nameless  
```{r processing shape_ids, message=FALSE, warning=FALSE,results = FALSE}
res <- map(1:nrow(bs_lines),function(.x){
  out <- tryCatch(
    {print(bs_lines[.x,]$shape_id)
      # filter relevant stops and intersections
      one <- joined %>% filter(shape_id == bs_lines[.x,]$shape_id)
      two <- joined_inter%>% filter(shape_id == bs_lines[.x,]$shape_id)
      # blend intersections to sfnetwork
      net <- bs_lines[.x,] %>% 
        as_sfnetwork(directed = T) %>% 
        convert(to_spatial_subdivision) %>%
        st_network_blend(two)
      # filter segments where there are stops
      rel_rows <- net %E>%
        st_as_sf() %>% 
        mutate(rn = row_number()) %>% 
        st_buffer(-10,singleSide = T) %>% 
        filter(apply(st_intersects(geometry,one,sparse = F),1,any)) 
      # get stops ids in order to blenf into network
      rel_idx <- net %E>%
        st_as_sf() %>% 
        mutate(rn = row_number()) %>% 
        st_buffer(-10,singleSide = T) %>% 
        st_intersects(one,sparse = F) %>% 
        apply( 1,which)
      # create sfnetwork for each segment
      nets <- net %E>%
        st_as_sf() %>% 
        mutate(rn = row_number(),
               idx = rel_idx) %>% 
        filter(rn %in% rel_rows$rn) %>% 
        mutate(
          # create an sfnetwork for each segment according to row number
          net1 = map2(geometry,rn, ~.x %>% 
                        st_sfc(crs = 2039) %>% 
                        st_sf() %>% 
                        as_sfnetwork(directed = T) %E>% 
                        mutate(rn = .y)),
          # blend stops into each sfnetwork
          net2 = map2(net1,idx, function(x,y){
            st_network_blend(x,one[y,]) %>% 
              mutate(rn2 = row_number())
          })) %>% 
        pull(net2)
      # get all joined net stops
      stops <- bind_rows(map(nets,~.x %N>% st_as_sf())) %>% 
        filter(!is.na(stop_id)) %>% 
        distinct()
      # bind all graphs together, including those segments without stops
      final_net <- net %E>%
        mutate(rn = row_number()) %>%
        filter(!rn %in% rel_rows$rn) %E>%
        bind_graphs(bind_graphs(nets)) %>%
        # convert to sfnetwork
        as_sfnetwork(edges_as_lines = T,directed = T) %E>% 
        st_as_sf() %>%
        # arrange by order
        arrange(rn,rn2) %>% 
        as_sfnetwork(directed = T)
      route <- final_net %E>% 
        st_as_sf() %>% 
        mutate(cs = cumsum(st_length(geometry))) 
      # create cuts
      cuts <- final_net %>% 
        st_join(stops) %>%
        st_join(two) %>% 
        mutate(rn = row_number()) %>% 
        as_tibble() %>% 
        as_tibble() %>%
        select(geometry,stop_id,inter_id,rn)
      # process route and cuts
      phase_1 <- route %>%
        # join all cuts
        left_join(cuts,by = c("from" = "rn")) %>%
        left_join(cuts,by = c("to" = "rn")) %>%
        # fill all shape_ids to prevent errors
        fill(shape_id,.direction = "downup") %>%
        # give each origin and destination an id: stop id if stop, intersection id if intersection, start\end if edges
        mutate(origin = ifelse(!is.na(stop_id.x),stop_id.x,ifelse(!is.na(inter_id.x),inter_id.x,NA)),
               destination = ifelse(!is.na(stop_id.y),stop_id.y,ifelse(!is.na(inter_id.y),inter_id.y,NA)),
               origin = ifelse(row_number() == 1,paste0("start_",shape_id),origin),
               destination = ifelse(row_number() == nrow(.),paste0("end_",shape_id),destination)) %>%
        select(origin,destination,from,to) %>%
        fill(origin) %>%
        fill(destination,.direction = "up") %>%
        # preserve order
        mutate(lag1 = lag(origin),
               lag2 = lag(destination),
               counter =coalesce(ifelse(origin == lag1 & destination == lag2,0,1),1) %>% cumsum()) %>% 
        select(-lag1,-lag2)
      # join created lines, some have no named edge
      phase_2 <- phase_1 %>%
        group_by(origin,destination,counter) %>%
        summarise(do_union=T)
      # prevent bugs if no multilinestring
      phase_21 <- phase_2 %>%
        filter(st_geometry_type(geometry.x) == "MULTILINESTRING")
      if(nrow(phase_21) == 0){
        phase_3 <- phase_2
      }else{
        phase_3 <- bind_rows(phase_2 %>%
                               filter(!st_geometry_type(geometry.x) == "MULTILINESTRING"),
                             phase_2 %>%
                               filter(st_geometry_type(geometry.x) == "MULTILINESTRING") %>%
                               st_line_merge())
      }
      phase_3 %>%
        ungroup() %>%
        arrange(counter) %>%
        # st_buffer(-10,singleSide = T) %>%
        mutate(rn3 = row_number())},
      error=function(cond) {
        message("Here's the original error message:")
        message(cond)
        # Choose a return value in case of error
        return(NA)},
        finally = print(paste("line number",.x))
      
  )
  return(out)
  
})
```

## map of products

```{r map of lines and cuts}
names(res) <- bs_lines$shape_id
resna <- res[which(is.na(res))]
resnotna <- res[!which(is.na(res))]
with_shapes <- map(names(resnotna),~resnotna[[.x]] %>% mutate(shape_id = .x)) %>% bind_rows()
starts <- with_shapes %>% st_startpoint() %>% st_sf()
# m1 <- mapview(with_shapes) + mapview(starts)
# m1@map
gc()
save.image("all_israel_pitstop.RData")
```

```{r cleanup}
rm(res)
rm(resnotna)
rm(spo_gtfs)
rm(joined_inter)
rm(starts)
gc()
```



## stops in routes - segmenting 1,2,3 order cuts
we remove the israel rail lines due to the fact they belong to a different network
stops_in_routes includes all of the stops in each shape_id  

prod1 classifies the cuts according to their order, the following way:  
for each shape_id:  
for each row:  
classify the order of the cut:  
first order - start,end, or stop in shape_id  
second order - stop not in shape_id  
third order - intersections  


prod2 classifies the segments according to their order, the following way:  
for each shape_id:  
for each row:  
if one of the cuts is third order - segment is third order  
if no cut is third order but one is second order - segment is second order  
if no cut is second ot third order - segment is first order  
```{r order cuts}
# remove missing routes because of algorithm
routes_to_remove <- names(resna)
stops_in_routes <- bs_gtfs$routes %>% 
  left_join(bs_gtfs$trips,by = "route_id") %>% 
  left_join(bs_gtfs$stop_times,by = "trip_id") %>% 
  select(route_id,route_desc,shape_id,stop_id,agency_id) %>% 
  filter(!shape_id %in% routes_to_remove) %>% 
  distinct()  
# remove israel rail - useless
to_remove <- stops_in_routes %>% filter(agency_id== 2) %>% distinct(shape_id) %>% pull(shape_id)
with_shapes <- with_shapes %>% filter(!shape_id %in% to_remove)

with_shapes <- with_shapes %>% 
  # for each shape_id, add cumulative sum of length to know distance in line
  mutate(length = st_length(geometry.x)) %>% 
  as_tibble() %>% 
  group_by(shape_id) %>% 
  mutate(length= as.numeric(length),
         cs = replace_na(lag(cumsum(length)),0)) %>% 
  left_join(stops_in_routes,by = c("origin" = "stop_id","shape_id")) %>% 
  left_join(stops_in_routes,by = c("destination" = "stop_id","shape_id"))
# getting relevent makatim without israel rail

makats <- bs_gtfs$routes %>% 
  filter(!agency_id== 2,route_id %in% unique(stops_in_routes$route_id) ) %>% 
  left_join(bs_gtfs$trips,by = "route_id") %>% 
  select(route_id,route_desc,shape_id) %>% 
  distinct()

prod1 <- makats %>% 
  left_join(with_shapes,by = "shape_id") %>%
  select(-route_id,-route_id.x,-route_id.y,-agency_id.x,-agency_id.y,-rn3) %>% 
  mutate(start = str_detect(origin, "start"),
         end = str_detect(destination, "end"),
         is_stop_in_orig = coalesce(route_desc == route_desc.x,FALSE),
         is_stop_in_dest = coalesce(route_desc == route_desc.y,FALSE),
         is_orig_intersection = str_detect(origin,"inter"),
         is_dest_intersection = str_detect(destination,"inter"),
         orig_cut = ifelse(start | is_stop_in_orig, "first",
                           ifelse(is_orig_intersection,"third","second")),
         dest_cut = ifelse(end | is_stop_in_dest, "first",
                           ifelse(is_dest_intersection,"third","second")))
prod2 <- prod1 %>% 
  select(-c(route_desc.x:is_dest_intersection)) %>% 
  mutate(segment_type = ifelse(orig_cut == "third" | dest_cut == "third", "third",
                               ifelse(orig_cut =="second"|dest_cut =="second","second","first")),
         counter_first = orig_cut == "first",
         counter_second = orig_cut == "second",
         counter_third = orig_cut == "third"
  ) %>% 
  group_by(shape_id) %>% 
  mutate(counter_seg_first = cumsum(counter_first)) %>% 
  ungroup() %>% 
  group_by(shape_id,counter_seg_first) %>% 
  mutate(counter_seg_second = cumsum(counter_second),
         has_second = sum(counter_second) > 0) %>% 
  ungroup() %>% 
  group_by(shape_id,counter_seg_first,counter_seg_second) %>% 
  mutate(counter_seg_third = cumsum(counter_third),
         has_third = sum(counter_third)>0) %>% 
  mutate(geometry.x = st_transform(geometry.x, 4326))
save.image("all_israel_pitstop2.RData")
```

## creating tables
taking all materials and processing them into 7 tables:  
cuts - all cuts ids, type (start, end, stop, intersection) and location  
routes - route_id,route_desc(line-direction-altenative),shape_id for relevant shapes  
segments - all unique geometries with uid, length and wkt geometry
first_order_cuts(should be changed to first_order_segments) - for each shape_id, all first order segments ids, and their start and end cuts
second_order_cuts(should be changed to second_order_segments) - for each shape_id, all second order segments ids by their containing first order segment id, and their start and end cuts
third_order_cuts(should be changed to second_order_segments) - for each shape_id, all third order segments ids by their containing first and second order segment id, and their start and end cuts
route order - for each shape id, the order of the segments, a reference to the unique segment id, and distance from the line origin
```{r to tables, message=FALSE, warning=FALSE,results = FALSE}
load("all_israel_pitstop2.RData")
rm(bs_gtfs)
rm(prod1)
rm(with_shapes)

final <- list()
final$routes <- makats %>% 
  mutate(route_id=as.integer(route_id),
         shape_id = shape_id) %>% 
  select(route_id,shape_id,route_desc)
st_x = function(x) round(st_coordinates(x)[,1],5)
st_y = function(x) round(st_coordinates(x)[,2],5)
final$cuts <- bind_rows(
  prod2 %>% ungroup() %>% 
    select(origin,geometry.x) %>% 
    mutate(geom = st_startpoint(geometry.x)) %>%
    select(-geometry.x) %>% 
    st_sf() %>% 
    rename(cut_id = origin),
  prod2 %>% 
    ungroup() %>% 
    select(destination,geometry.x) %>% 
    mutate(geom = st_endpoint(geometry.x)) %>%
    select(-geometry.x) %>% 
    st_sf() %>% 
    rename(cut_id = destination)
) %>% distinct() %>% 
  mutate(intersection = str_detect(cut_id,"inter"),
         start = str_detect(cut_id,"start"),
         end = str_detect(cut_id,"end"),
         type = ifelse(intersection,"inter",ifelse(start,"start",ifelse(end,"end","stop"))),
         lon = st_x(geom),
         lat = st_y(geom)) %>% 
  st_drop_geometry() %>% 
  select(cut_id,type,lon,lat)

final$segments <- prod2 %>% 
  ungroup() %>% 
  select(geometry = geometry.x,length) %>% 
  distinct() %>% 
  mutate(seg_UID = row_number(),
         geometry = map_chr(geometry,st_as_text)) %>% 
  select(seg_UID,length,geometry)


tmp <- prod2 %>%
  mutate(geometry.x = map_chr(geometry.x,st_as_text)) %>% 
  left_join(final$segments, by = c("geometry.x"="geometry"))
rm(prod2)
final$first_order_cuts <- tmp %>% 
  ungroup() %>% 
  select(shape_id,
         first_order_seg_id = counter_seg_first,
         from_cut = origin,
         to_cut = destination) %>% 
  group_by(shape_id,first_order_seg_id) %>% 
  mutate(from_cut = first(from_cut),
         to_cut = last(to_cut)
  ) %>% 
  distinct()
final$second_order_cuts <- tmp %>% 
  ungroup() %>% 
  filter(has_second) %>% 
  select(shape_id,
         first_order_seg_id = counter_seg_first,
         second_order_seg_id = counter_seg_second,
         from_cut = origin,
         to_cut = destination) %>% 
  distinct() %>% 
  group_by(shape_id,first_order_seg_id,second_order_seg_id) %>% 
  mutate(from_cut = first(from_cut),
         to_cut = last(to_cut)
  ) %>% 
  distinct()  
final$third_order_cuts <- tmp %>% 
  ungroup() %>% 
  filter(has_third) %>% 
  select(shape_id,
         first_order_seg_id = counter_seg_first,
         second_order_seg_id = counter_seg_second,
         third_order_seg_id = counter_seg_third,
         from_cut = origin,
         to_cut = destination,
         has_second) %>% 
  mutate(second_order_seg_id = ifelse(has_second,second_order_seg_id,NA_integer_)) %>% 
  select(-has_second) 
final$route_order <- tmp %>% 
  ungroup() %>% 
  select(shape_id, 
         serial_cut_number = counter, 
         first_order_seg_id= counter_seg_first,
         second_order_seg_id= counter_seg_second,
         third_order_seg_id= counter_seg_third,
         seg_UID,
         distance_from_origin = cs,
         serial_cut_number =counter,
         has_second,
         has_third) %>% 
  mutate(second_order_seg_id = ifelse(has_second,second_order_seg_id,NA_integer_),
         third_order_seg_id = ifelse(has_third,third_order_seg_id,NA_integer_)) %>% 
  select(-has_second,-has_third)
map(names(final),~write_csv(final[[.x]],paste0(.x,".csv")))
```

## creating a data model of the tables, including all foreign keys

```{r data model, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results=false}
new_model <- df_to_model(
  list=names(final), envir=as.environment(final)
)
new_model %>% 
  add_foreign_key("route_order","seg_UID","segments","seg_UID") %>%
  add_foreign_key("first_order_cuts","from_cut","cuts","cut_id") %>% 
  add_foreign_key("first_order_cuts","to_cut","cuts","cut_id") %>% 
  add_foreign_key("second_order_cuts","from_cut","cuts","cut_id") %>% 
  add_foreign_key("second_order_cuts","to_cut","cuts","cut_id") %>%
  add_foreign_key("third_order_cuts","from_cut","cuts","cut_id") %>% 
  add_foreign_key("third_order_cuts","to_cut","cuts","cut_id") %>% 
  add_foreign_key("first_order_cuts","first_order_seg_id","route_order","first_order_seg_id") %>% 
  add_foreign_key("second_order_cuts",c("first_order_seg_id","second_order_seg_id"),"route_order",c("first_order_seg_id","second_order_seg_id")) %>% 
  add_foreign_key("third_order_cuts",c("first_order_seg_id","second_order_seg_id","third_order_seg_id"),"route_order",c("first_order_seg_id","second_order_seg_id","third_order_seg_id")) %>% 
  add_foreign_key("route_order","shape_id","routes","shape_id") %>%
  add_foreign_key("first_order_cuts","shape_id","routes","shape_id") %>%
  add_foreign_key("second_order_cuts","shape_id","routes","shape_id") %>%
  add_foreign_key("third_order_cuts","shape_id","routes","shape_id") %>%
  model_relational_data()
```

